---
title: "Zika Shannon  Diversity"
author: "Kate Johnson"
date: "May 20, 2019 - October 8, 2019 "
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r LoadLib, include=FALSE,message=F}
library(dplyr)
library(ggplot2)
library(tidyr)
library(FSA)
library(rcompanion)
```

```{r LoadData,echo=FALSE,message=F}
#load and rearrange data if needed
cov = 200
freq =0.03
filename=paste0('../../Diversity/PCR1/ShannonDiversity_STAR.zika.perkb.',cov,'.',freq,'.pcr1.csv') 
mydata<-read.csv(file=filename,header=T,sep=",",na.strings = c('')) 
mydata = filter(mydata, segment=='FULL') %>% droplevels()
metadata = '../../MetadataZika/Zika_Metadata_v4.csv' #where metadata is located
meta_df = read.csv(file=metadata,header=T,sep=",",na.strings = c('nan')) %>%
  select(name,full,experiment)
mydata_meta = merge(mydata, meta_df,by.x='sample',by.y='name')
mydata=mydata_meta

tissue_list=c('brain','liver','kidney','spinal_cord','bone_marrow')

tissue_list2 = c('brain','liver','kidney','spinal_cord','bone_marrow',
                 'stock2','mosquito','spleen',
                 'stock1','blood','reproductive')

No_Stock = filter(mydata, sample!='stock' & 
                    tissue %in% tissue_list & segment=='FULL')%>%
  droplevels()

MouseOrder = c('M1','M3','M2','M4','M5','M6','M7','M8','M9','M10')
No_Stock$mouse = factor(No_Stock$mouse,levels=rev(MouseOrder))
```

```{r GeneDF,echo=FALSE,message=F}
Gene_ID = c('C','pr','M','E','NS1','NS2A','NS2B','NS3','NS4A','2K','NS4B','NS5')
Gene_start = c(107,473,752,977,2477,3533,4211,4601,6452,6833,6902,7655)
Gene_End = c(472,751,976,2476,3532,4210,4600,6451,6832,6901,7654,10363) #ns3 ns4a off by 100 missing
Gene_df = cbind(Gene_ID,Gene_start,Gene_End)
Gene_df = as.data.frame(Gene_df)
Gene_df$Gene_start = (as.numeric(as.character(Gene_df$Gene_start)))-106
Gene_df$Gene_End = as.numeric(as.character(Gene_df$Gene_End))-106
```

```{r tissueColors, echo=FALSE,message=F}
myColors<-c('black','#a9a9a9','#004d99','#d80000','#ee7600')
names(myColors) =c('brain','liver','kidney','spinal_cord','bone_marrow')
colScale <- scale_color_manual(name = "grp",values = myColors)

reducedColors =c('#66c2a4','#fc8d62','#8d9fcb','#e78ac3',
                 '#a6d854','#ffd92f','#e5c494','#b2b3b2',
                 '#feffb3','#0a068b','#fe3636')

names(reducedColors) = c('brain','liver','kidney','spinal_cord','bone_marrow',
                 'stock2','mosquito','spleen','stock1','blood','reproductive')

RedColScale=scale_color_manual(name='grp',values=reducedColors)

RedColFill = scale_fill_manual(name='grp',values=reducedColors)
```
# PCR1 ONLY

### Shannon Entropy broken up by tissue type, trianges represent samples from the second experiment.     
```{r, echo=F,message=F}
g=ggplot(data = filter(mydata,tissue %in% tissue_list2), aes(x = tissue, y = shannon_sum,
                                                             group=tissue))+
  geom_boxplot(width=0.5) +
  geom_jitter(aes(color=tissue,shape=experiment),
              size=0.95,alpha=1,width=0.2)+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=5)) +
  RedColScale
ggsave(g, file=("../../Diversity/PCR1/Zika.Tissue.Reduced.pdf"), width=5, height=3,limitsize=FALSE)
print(g) 
```

### PCR1 diversity measurements broken up by transmission groups.     
```{r, echo=F,message=F}
b<- ggplot(mydata, aes(x = transmission, y = shannon_sum,group=transmission))+
    geom_boxplot(width=0.5) +
    geom_jitter(aes(shape=experiment),width = 0.15,alpha=0.9)+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    theme_bw()
ggsave(b, file=("../../Diversity/PCR1/Zika.Transmission.PCR1.pdf"), width=6, height=4,limitsize=FALSE)
print(b) 
```

### Kruskal Wallis test to determine significance between transmission groups.    

Alternative hypothesis accepted- samples come from different populations. P.value is significant.    

P-value:  
```{r StatsTransmissionPCR1,echo=F,message=F,message=F}
transstat = filter(mydata,tissue %in% tissue_list2) %>%
  select(mouse,tissue,transmission,shannon_sum)


#summarize
s = Summarize(shannon_sum~transmission,data=transstat,digits=3)
write.csv(s,"../../Diversity/PCR1/SummaryStats.Diversity.ByTransmission.pcr1.csv",
          row.names=F)

#kruskal wallis test
pval=kruskal.test(shannon_sum~transmission,data=transstat)
print(pval$p.value)
#pval 0.009404
#Null: all samples/groups are from identical populations
#Alt: At least one of the samples comes from different population
#there exists a significant difference among the groups p.value = 0.009404

#Dunns test
#If the Kruskal–Wallis test is significant, a post-hoc analysis can be performed to #determine which groups differ from each other group. 
DT = dunnTest(shannon_sum~transmission,data=transstat,method = 'bh')
d = as.data.frame(DT$res)
write.csv(d,"../../Diversity/PCR1/DunnTest.Stats.Diversity.ByTransmission.PCR1.csv",row.names = F)
```

### Shannon entropy by mouse, and tissue for PCR1.    
```{r, echo=F,message=F,warning=F}
MLineAll = ggplot(data=filter(mydata,tissue%in%tissue_list2),
                  aes(x=mouse,y=shannon_sum,group=mouse,color=tissue)) +
  geom_line(color='black') +
  geom_point(aes(color=tissue,shape=experiment),alpha=0.9,size=2) +
  theme_bw()+
  coord_flip()+
  RedColScale
print(MLineAll)
ggsave(MLineAll, file=("../../Diversity/PCR1/Mouse.Tissue.Line.All.pdf"), width=7, 
       height=4,limitsize=FALSE,useDingbats=F)
```


```{r meanpcr1, echo=F,message=F}
mean.it= aggregate(data=mydata, shannon_sum ~ tissue +transmission ,FUN=mean)
sd.it = aggregate(data=mydata, shannon_sum ~ tissue +transmission ,FUN=sd)
colnames(sd.it) = c('tissue','transmission','sd')
mean.sd = merge(mean.it,sd.it,by=c('tissue','transmission'))
mean.sd$maxy = mean.sd$shannon_sum + mean.sd$sd
mean.sd$miny = mean.sd$shannon_sum - mean.sd$sd
mean.sd$tissue_transmission = paste0(mean.sd$transmission,'-',mean.sd$tissue)
mydata$tissue_transmission = paste0(mydata$transmission,'-',mydata$tissue)
```

### Bar graphs of the mean and standard deviation, using the shannon entropy of PCR1 samples.     

```{r meanplotpcr1, echo=F,message=F,warning=F}
mean.sd = filter(mean.sd, tissue %in% tissue_list2)

tt_ord = c('stock1-stock1','stock2-stock2','blood_meal-mosquito',
           'mosquito-brain','mosquito-spinal_cord','mosquito-kidney',
           'mosquito-liver','mosquito-spleen','mosquito-bone_marrow','mosquito-blood',
           'mosquito-reproductive','needle-brain','needle-spinal_cord',
           'needle-kidney','needle-liver','needle-spleen','needle-bone_marrow',
           'needle-blood','needle-reproductive')

mean.sd$tissue_transmission = factor(mean.sd$tissue_transmission,levels=tt_ord)
mydata = filter(mydata, tissue %in% tissue_list2)
#h=ggplot(data=filter(mean.sd, tissue%in% tissue_list2),
h=ggplot(data=mean.sd, aes(x=tissue_transmission, y=shannon_sum,
                           group=tissue_transmission)) + 
  geom_col(aes(fill=tissue))+
  geom_errorbar(aes(ymin=miny, ymax=maxy), width=.2,
                position=position_dodge(0)) +
  geom_jitter(data =filter(mydata,tissue %in% tissue_list2),
              aes(y=shannon_sum,x=tissue_transmission,group=tissue_transmission,
                  shape=experiment),
              width=0.2) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  RedColFill
ggsave(h, file= "../../Diversity/PCR1/Shannon.Mean.Tissue.pdf", 
       width=8, height=5,limitsize=FALSE, useDingbats=FALSE)
print(h)
```

### Kruskal Wallis statistics of shannon entropy by transmission and tissue using PCR1 samples.    
     
Alternative hypothesis accepted- samples by tissue type and transmission route are significantly different.    
      
p-value:     

```{r ShannonStats, echo=F,message=F,warning=F}
df = filter(mydata,tissue %in% tissue_list2) %>%
  select(mouse,tissue,transmission,shannon_sum,tissue_transmission)


#summarize
s = Summarize(shannon_sum~tissue_transmission,data=df,digits=3)
write.csv(s,"../../Diversity/PCR1/SummaryStats.Diversity.pcr1.csv",
          row.names=F)

#kruskal wallis test
pval=kruskal.test(shannon_sum~tissue_transmission,data=df)
print(pval$p.value)
#pval 0.004
#Null: all samples/groups are from identical populations
#Alt: At least one of the samples comes from different population
#there exists a significant difference among the groups p = 0.004207

#Dunns test
#If the Kruskal–Wallis test is significant, a post-hoc analysis can be performed to #determine which groups differ from each other group. 
DT = dunnTest(shannon_sum~tissue_transmission,data=df,method = 'bh')
d = as.data.frame(DT$res)
write.csv(d,"../../Diversity/PCR1/DunnTest.Stats.Diversity.PCR1.csv",row.names = F)
```

# FULL LENGTH
     
### Shannon entropy by tissue type across the full-length of the genome.     
```{r FullGenomeShannon,echo=F,warning=F}
full<-paste0('../../Diversity/FullLength/ShannonDiversity_STAR.zika.ntpos.',
             cov,'.',freq,'.full.csv') 

ntpos_df = read.csv(file=full,header=T,sep=",",na.strings = c(''))
ntpos_df = filter(ntpos_df, tissue %in% tissue_list2)%>%
  droplevels()

shannon_ntpos = ggplot(ntpos_df)+ 
  geom_bar(aes(x=ntpos,y=shannon,group=tissue,fill=tissue),width=50,
           stat='identity',alpha=1) +
  theme_bw() +
  RedColFill
ggsave(shannon_ntpos, file=("../../Diversity/FullLength/ntpos.shannon.bar.thick.pdf"), 
       width=10, height=3,limitsize=FALSE)

print(shannon_ntpos)
```


```{r LoadFull,echo=F,message=F}
filename=paste0('../../Diversity/FullLength/ShannonDiversity_STAR.zika.perkb.',cov,'.',freq,'.full.csv') 
mydata<-read.csv(file=filename,header=T,sep=",",na.strings = c('')) 
mydata = filter(mydata, segment=='FULL' & tissue %in% tissue_list2) %>% 
  droplevels()
mydata_meta = merge(mydata, meta_df,by.x='sample',by.y='name')
mydata=mydata_meta %>% filter(full =='yes') %>% droplevels()

```


```{r meanfull, echo=F,message=F}
mean.it= aggregate(data=mydata, shannon_sum ~ tissue +transmission ,FUN=mean)
sd.it = aggregate(data=mydata, shannon_sum ~ tissue +transmission ,FUN=sd)
colnames(sd.it) = c('tissue','transmission','sd')
mean.sd = merge(mean.it,sd.it,by=c('tissue','transmission'))
mean.sd$maxy = mean.sd$shannon_sum + mean.sd$sd
mean.sd$miny = mean.sd$shannon_sum - mean.sd$sd
mean.sd$tissue_transmission = paste0(mean.sd$transmission,'-',mean.sd$tissue)
mydata$tissue_transmission = paste0(mydata$transmission,'-',mydata$tissue)
```

### Mean and standard deviation using the Shannon entropy from samples where the entire genome was sequenced.     
```{r meanplotfull, echo=F,message=F,warning=F}
mean.sd = filter(mean.sd, tissue %in% tissue_list2)%>%
  droplevels()

tt_ord = c('stock1-stock1','stock2-stock2','blood_meal-mosquito',
           'mosquito-brain','mosquito-spinal_cord','mosquito-kidney',
           'mosquito-liver','mosquito-spleen','mosquito-bone_marrow','mosquito-blood',
           'mosquito-reproductive','needle-brain','needle-spinal_cord',
           'needle-kidney','needle-liver','needle-spleen','needle-bone_marrow',
           'needle-blood','needle-reproductive')

mean.sd$tissue_transmission = factor(mean.sd$tissue_transmission,levels=tt_ord)


h=ggplot(data=mean.sd, aes(x=tissue_transmission, y=shannon_sum,
                           group=tissue_transmission)) + 
  geom_col(aes(fill=tissue))+
  geom_errorbar(aes(ymin=miny, ymax=maxy), width=.2,
                position=position_dodge(0)) +
  geom_jitter(data=mydata,
              aes(y=shannon_sum,x=tissue_transmission,group=tissue_transmission,
                  shape=experiment),
              width=0.2) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  RedColFill
ggsave(h, file= "../../Diversity/FullLength/Shannon.Mean.Tissue.FULL.pdf", 
       width=8, height=5,limitsize=FALSE, useDingbats=FALSE)
print(h)
```

### Kruskal-Wallis test on the tissue type and transmission route from samples that have the full-length genome.     
    
p-value:     
```{r FullShannonStats,echo=F,message=F,warning=F}
df = filter(mydata,tissue %in% tissue_list2) %>%
  select(mouse,tissue,transmission,shannon_sum,tissue_transmission)


#summarize
s = Summarize(shannon_sum~tissue_transmission,data=df,digits=3)
write.csv(s,"../../Diversity/FullLength/SummaryStats.Diversity.Full.csv",
          row.names=F)

#kruskal wallis test
pval=kruskal.test(shannon_sum~tissue_transmission,data=df)
print(pval$p.value)
#pval 0.004
#Null: all samples/groups are from identical populations
#Alt: At least one of the samples comes from different population
#there exists a significant difference among the groups p = 0.004207

#Dunns test
#If the Kruskal–Wallis test is significant, a post-hoc analysis can be performed to #determine which groups differ from each other group. 
DT = dunnTest(shannon_sum~tissue_transmission,data=df,method = 'bh')
d = as.data.frame(DT$res)
write.csv(d,"../../Diversity/FullLength/DunnTest.Stats.Diversity.FULL.csv",row.names = F)
```

Boxplots of Shannon Entropy broken up by transmission route.     
     
Uses only samples that have the full genome.     

```{r, echo=F,message=F}
b<- ggplot(mydata, aes(x = transmission, y = shannon_sum,group=transmission))+
    geom_boxplot(width=0.5) +
    geom_jitter(aes(shape=experiment),width = 0.15,alpha=0.9)+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    theme_bw()
ggsave(b, file=("../../Diversity/FullLength/Zika.Transmission.FULL.pdf"), width=6, height=4,limitsize=FALSE)
print(b) 
```

### Kruskal Wallis statistics on full-length samples grouped by transmission route.     
      
Not significant.     
       
P-value:    
       
```{r StatsTransmissionFULL,echo=F,message=F,warning=F}
transstat = filter(mydata,tissue %in% tissue_list2) %>%
  select(mouse,tissue,transmission,shannon_sum)


#summarize
s = Summarize(shannon_sum~transmission,data=transstat,digits=3)
write.csv(s,
          "../../Diversity/FullLength/SummaryStats.Diversity.ByTransmission.FULL.csv",
          row.names=F)

#kruskal wallis test
pval=kruskal.test(shannon_sum~transmission,data=transstat)
print(pval$p.value)
#pval 0.1577169
#Null: all samples/groups are from identical populations
#Alt: At least one of the samples comes from different population
#there does not exist a significant difference among the groups p.value 0.1577169

#Dunns test
#If the Kruskal–Wallis test is significant, a post-hoc analysis can be performed to #determine which groups differ from each other group. 
DT = dunnTest(shannon_sum~transmission,data=transstat,method = 'bh')
d = as.data.frame(DT$res)
write.csv(d,"../../Diversity/FullLength/DunnTest.Stats.Diversity.ByTransmission.FULL.csv",row.names = F)
```










